---
title: "Plant Microbiome Metaanalysis"
output: html_notebook
author: "Joe Edwards"
---

## Introduction

## Load in the libraries
Note that this analysis relies on the tidyMB package. This can be found on my github profile.
```{r}
library(vegan)
library(broom)
library(tidyverse)
library(broom)
library(biobroom)
library(tidyMB)
library(DESeq2)
```

```{r}
## Do Not Run
# wExp <- read_rds("~/RMB/SoilDomesticationData/wExp.rds")
# wExp %>% 
  # filter(!Experiment%in%c("Mix", "Transplant", "Weeds")) %>% 
  # ungroup() %>% 
  # group_by(variable) %>% 
  # filter(sum(value) > 0) %>% 
  # dplyr::select(-From, - Site, -LatinName, -host_subject_id) %>% 
  # filter(!Compartment%in%c("fresh water", "Grapes", "Leaves", "Rhizoplane"))
  # saveRDS("~/RMB/SoilDomestication/Data/ma_data.rds")

mExp <- readRDS("~/RMB/SoilDomesticationData/ma_data.rds") %>% filter(!Compartment%in%c("fresh water", "Grapes", "Leaves", "Rhizoplane"))
tax <- readRDS("~/RMB/SoilDomestication/Data/gg_otus_tax.rds")
```

```{r}
plant_cap <- tidy_cap(mExp %>% ungroup() %>% mutate(logRA = log2(RA * 1000 + 1)), value = "logRA", formula = "host_common_name")
plant_pcoa <-  tidy_pcoa(mExp %>% ungroup() %>% mutate(logRA = log2(RA * 1000 + 1)), value = "logRA")

write_rds(plant_cap, "~/RMB/SoilDomestication/Data/mPlant_cap.rds")
write_rds(plant_pcoa, "~/RMB/SoilDomestication/Data/mPlant_pcoa.rds")
```

```{r}
plant_cap$axes %>% 
  ggplot(aes(CAP1, CAP2, color = study)) +
  geom_point(alpha = 0.5) +
  scale_color_brewer(palette = "Set2") +
  theme_minimal()

plant_pcoa$axes %>% 
  ggplot(aes(MDS1, MDS2, color = study)) +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = brewer.pal(7, "Set2")[-4]) +
  labs(x = paste("PCo1 (", round(plant_pcoa$eigen_vals[1] * 100, 2), "%)", sep = ""), 
       y = paste("PCo2 (", round(plant_pcoa$eigen_vals[2] * 100, 2), "%)", sep = "")) +
  theme_minimal()
```
```{r}
mAdon <- long_adonis(mExp %>% ungroup() %>% mutate(logRA = log2(RA * 1000 + 1)), value = "logRA", formula = "Compartment + study")
mAdon %>% tidy()
```

## Alpha diversity
Honestly, I hate this part. I really don't see the need for this analysis. What does it mean? What does it matter, espcieally given that different metrics give different results? Sigh... Here we go anyway. I guess.
```{r}
mAdiv <- mExp %>% 
  filter(RA > 0) %>% 
  ungroup() %>% 
  group_by(Compartment, host_common_name, study, SampleID, depth) %>% 
  summarise(richness = sum(RA > 0), shannon = -sum(log2(RA) * RA))
```

Let's plot out the results
```{r}
mAdiv %>% ungroup() %>% 
  mutate(Compartment = ifelse(grepl("soil", Compartment, ignore.case = T), "Bulk Soil", Compartment)) %>% 
  mutate(Compartment = ifelse(grepl("rhizosphere", Compartment, ignore.case = T), "Rhizosphere", Compartment)) %>% 
  mutate(Compartment = ifelse(grepl("root", Compartment, ignore.case = T), "Endosphere", Compartment)) %>% 
  mutate(Compartment = fct_relevel(factor(Compartment), "Bulk Soil", "Rhizosphere", "Endosphere")) %>% 
  ggplot(aes(study, shannon, color = study, group = Compartment)) +
  geom_point(position = position_jitterdodge(jitter.width = 0.1, jitter.height = 0), alpha = 0.5) +
  scale_color_manual(values = brewer.pal(7, "Set2")[-4]) +
  theme_minimal() +
  labs(x = "") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank(), legend.position = "none")

tidy(aov(shannon ~ study + Compartment, mAdiv))
```

```{r}
mAdiv %>% ungroup() %>% 
  mutate(Compartment = ifelse(grepl("soil", Compartment, ignore.case = T), "Bulk Soil", Compartment)) %>% 
  mutate(Compartment = ifelse(grepl("rhizosphere", Compartment, ignore.case = T), "Rhizosphere", Compartment)) %>% 
  mutate(Compartment = ifelse(grepl("root", Compartment, ignore.case = T), "Endosphere", Compartment)) %>% 
  mutate(Compartment = fct_relevel(factor(Compartment), "Bulk Soil", "Rhizosphere", "Endosphere")) %>% 
  ggplot(aes(depth, shannon, color = study, group = Compartment)) +
  geom_point(alpha = 0.5) +
  facet_grid(.~Compartment) +
  scale_color_manual(values = brewer.pal(7, "Set2")[-4]) +
  scale_x_log10() +
  theme_minimal() +
  theme(legend.position = "none")

mAdiv %>% 
  group_by(Compartment) %>% nest() %>% 
  mutate(lm_mod = map(data, ~lm(shannon ~ depth, .))) %>% 
  unnest(map(lm_mod, ~tidy(.)))
```

```{r}
mDist <- mExp %>% 
  ungroup() %>% 
  mutate(Compartment = ifelse(grepl("soil", Compartment, ignore.case = T), "Bulk Soil", Compartment)) %>% 
  mutate(Compartment = ifelse(grepl("rhizosphere", Compartment, ignore.case = T), "Rhizosphere", Compartment)) %>% 
  mutate(Compartment = ifelse(grepl("root", Compartment, ignore.case = T), "Endosphere", Compartment)) %>% 
  filter(Compartment != "Bulk Soil") %>% 
  mutate(logRA = log2(RA * 1000 + 1)) %>% 
  long_distance(value = "logRA", keep_metadata = T)

mDist %>% 
  filter(study.x == "rice_edwards" | study.y == "rice_edwards") %>% 
  filter(study.x != study.y) %>% 
  mutate(study = ifelse(study.x == "rice_edwards", study.y, study.x)) %>% 
  mutate(study = fct_reorder(study, value)) %>% 
  ggplot(aes(study, value, color = study, alpha = study)) +
  geom_boxplot(alpha = 0, width = 0.3) +
  theme_minimal() +
  scale_color_manual(values = c("#E5C494", "#66C2A5", "#8DA0CB", "#A6D854", "#FC8D62", "#E5C494")) +
  labs(x = "") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  scale_alpha_manual(values = c(0.05, 0.05, 0.05, 0.05, 1))

mDist %>% 
  filter(study.x == "rice_edwards" | study.y == "rice_edwards") %>% 
  filter(study.x != study.y) %>% 
  mutate(study = ifelse(study.x == "rice_edwards", study.y, study.x)) %>% 
  mutate(study = fct_reorder(study, value)) %>% 
  lm(value ~ study, .) %>% 
  tidy(.)
```
```{r}
m_mean_dist <- mDist %>% 
  rowwise() %>% 
  mutate(study1 = min(study.x, study.y), study2 = max(study.x, study.y)) %>% 
  ungroup() %>% 
  group_by(study1, study2) %>% 
  summarise(mean_dis = mean(value))

m_mean_dist %>% filter(study1 != study2) %>% 
  ggplot(aes(study1, study2, fill = mean_dis, label = round(mean_dis, digits = 2))) +
  geom_tile(color = "black") +
  geom_text() +
  scale_fill_gradient(low = "white", high = 'black') 
```


## Phylum level analysis
Let's aggregate the counts at the phylum level and give it a go.
```{r}
mPhy <- mExp %>% 
  inner_join(tax, by = "variable") %>% 
  group_by(SampleID, Compartment, Phylum2, host_common_name, study) %>% 
  summarise(sample_total = sum(RA)) %>% ungroup() %>% 
  mutate(Compartment = ifelse(grepl("soil", Compartment, ignore.case = T), "Bulk Soil", Compartment)) %>% 
  mutate(Compartment = ifelse(grepl("rhizosphere", Compartment, ignore.case = T), "Rhizosphere", Compartment)) %>% 
  mutate(Compartment = ifelse(grepl("root", Compartment, ignore.case = T), "Endosphere", Compartment)) %>% 
  filter(!Compartment%in%c("fresh water", "Grapes", "Leaves", "Rhizoplane"))
```

Plot out the results
```{r}
mPhy %>% 
  group_by(Phylum2) %>% 
  mutate(phy_total = sum(sample_total)) %>% 
  group_by(Phylum2, phy_total) %>% 
  nest() %>% 
  arrange(-phy_total) %>% 
  head(15) %>% 
  unnest() %>% ungroup() %>% 
  group_by(Compartment, host_common_name, SampleID, study) %>% 
  nest() %>% 
  group_by(study, Compartment) %>% 
  mutate(gamma_sum = map_dbl(data, ~sum(.x %>% filter(Phylum2 == "Gammaproteobacteria") %>% pull(sample_total)))) %>% 
  arrange(gamma_sum) %>% 
  mutate(order = 1:n()) %>%
  unnest() %>% ungroup() %>% 
  mutate(Compartment = fct_relevel(factor(Compartment), "Bulk Soil", "Rhizosphere", "Endosphere")) %>% 
  ggplot(aes(order, sample_total, fill = Phylum2)) +
  geom_bar(stat = "identity", width = 1, position = "stack") +
  facet_grid(.~study + Compartment, scales = "free_x", space = "free") +
  scale_fill_manual(values = c(brewer.pal(11, "RdGy"), brewer.pal(5, "Blues")[-1])) +
  theme_minimal()
```

Next, let's ask the question: which phyla are consistently enriched or depleted in the rice studies compared to the other studies. The way we will go about doing this is to fit each OTU to two different models. One where the rice_edwards dataset is the reference and the other with rice_knight being the reference. Then find the phyla that behave consistently between the two models.
```{r}
rice_ed <- mPhy %>% 
  mutate(study = relevel(factor(study), ref = "rice_edwards")) %>% 
  filter(study != "Rice_knight") %>% 
  group_by(Phylum2) %>% 
  nest() %>% 
  mutate(model = map(data, ~lm(log2(sample_total * 1000 + 1) ~ study, .))) %>% 
  unnest(map(model, ~tidy(.))) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(p.adj = p.adjust(p.value, "BH")) %>% 
  filter(p.adj <= 0.05) %>% 
  mutate(direction = ifelse(estimate > 0, "other", "rice")) %>% 
  mutate(model = "rice_ed")

rice_kn <- mPhy %>% 
  mutate(study = relevel(factor(study), ref = "Rice_knight")) %>% 
  filter(study != "rice_edwards") %>% 
  group_by(Phylum2) %>% 
  nest() %>% 
  mutate(model = map(data, ~lm(log2(sample_total * 1000 + 1) ~ study, .))) %>% 
  unnest(map(model, ~tidy(.))) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(p.adj = p.adjust(p.value, "BH")) %>% 
  filter(p.adj <= 0.05) %>% 
  mutate(direction = ifelse(estimate > 0, "other", "rice")) %>% 
  mutate(model = "rice_kn")
```
Now that we have the two models, let's stick em together, filter out the phyla that are consistent between the models, and then identify the phyla that are enriched or depleted in rice compared to every other dataset.
```{r}
bind_rows(rice_ed, rice_kn) %>% 
  dplyr::count(direction, Phylum2, term) %>% 
  filter(n == 2) %>% 
  dplyr::count(direction, Phylum2) %>% arrange(-nn) %>% 
  filter(nn == 4) %>% 
  inner_join(mPhy %>% 
               group_by(Phylum2) %>% 
               summarise(total = sum(sample_total)) %>% 
               dplyr::select(Phylum2, total), by = "Phylum2") %>% 
  arrange(-total)
```

