---
title: "vigorMB_beta"
output: html_notebook
---

```{r}
source("~/RMB/SoilDomestication/VIGOR/General/rmb_functions.R")
library(DESeq2)
library(ggdendro)
library(tidyverse)
library(phyloseq)
library(cowplot)

soil.pal <- c("#92324B","#9e66a7",'#ee5768', '#fdc248', '#1ca04f','#27afb9')
phy.pal <- c("black", RColorBrewer::brewer.pal(11, "Spectral")[c(9,2,11)], RColorBrewer::brewer.pal(6, "Greys")[-1])

otu <- readRDS("~/RMB/SoilDomestication/VIGOR/Data/formatted_otu_table.RDS")
map <- readRDS("~/RMB/SoilDomestication/VIGOR/Data/formatted_map.RDS")
tax <- read_tax("~/RMB/SoilDomestication/VIGOR/General/gg_13_8_97_tax.tsv")
```

Get the top phylum/classes and collapse the rest into the "other" category
```{r}
otu.ra <- rel_ab(otu)
otu.ra.tidy <- tidy_otu(as.data.frame(otu.ra))

tax <- expand_proteo(tax)
top.tax <- get_top_taxa(otu.ra, tax, rank = "PhyClass", n = 7)
tax <- collapse_other(tax, top.tax)
```

Calculate beta diversity and PCos. I am using log transformed relative abundances because I couldn't use DESeq2 to normalize the whole dataset. This was because there is no single OTU with at least one count in all the samples, a requirement to run the default pipeline. 
```{r}
dist <- vegdist(log2(t(otu.ra) + 1))

pcoa.axes <- pcoa_axes(as.matrix(dist), map)
pcoa.eigval <- pcoa_eigval(as.matrix(dist), map)

ggplot(pcoa.axes, aes(Axis.1, Axis.2, color = Soil2, shape = paste(Study, Inoculum, sep = " / "))) +
  geom_point(size = 3, stroke = 2) +
  scale_shape_manual(values = c(2,17,1,16), name = "Study / Treatment") +
  scale_color_manual(values = soil.pal, name = "Soil") +
  xlab(paste("PCo1 (", pcoa.eigval$Eigval[1], "%)", sep = "")) +
  ylab(paste("PCo2 (", pcoa.eigval$Eigval[2], "%)", sep = "")) +
  #theme_minimal() +
  theme(text = element_text(size = 12),
        panel.background = element_rect(fill = "white"),
        panel.border = element_rect(fill = NA),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank())
```


Plot the dendogram and taxonomic profiles
```{r}
dd.row <- as.dendrogram(hclust(as.dist(dist), method = "average"))
ddata_x <- dendro_data(dd.row)

dendro.p <- ggplot(segment(ddata_x)) +
  geom_segment(aes(x=x, y=-y, xend=xend, yend=-yend))

labs <- label(ddata_x) %>% 
  inner_join(map, by = c("label" = "SampleID"))

phy.ra <- otu.ra.tidy %>% 
  inner_join(tax) %>% 
  group_by(SampleID, PhyClass2) %>%
  summarise(Total = sum(Count)) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = (Total/sum(Total))) %>% 
  inner_join(map, by = "SampleID") %>% 
  select(SampleID, PhyClass2, RelAb)

phy.ra <- inner_join(labs, phy.ra,  by = c("label" = "SampleID"))

dendro.p + 
  geom_bar(data = phy.ra, 
           aes(x=x, y=RelAb, fill = PhyClass2), stat = "identity") +
  geom_point(data=labs,
             aes(x=x, y=-.025, colour=Soil2, shape = paste(Study, Inoculum, sep = " / ")), size = 4, stroke = 1) +
  scale_color_manual(values = soil.pal, name = "Soil") +
  scale_fill_manual(values = phy.pal) +
  #scale_shape_manual(values = c(16,1)) +
  scale_shape_manual(values = c(2,17,1,16), name = "Study / Treatment") +
  coord_flip(expand = T) +
  theme_minimal() +
  theme(text = element_text(size = 15))
```

Visaulize them separately. In these analyses I used variance stabilizing transformed counts as the clustering with this approach was neater
First experiment
```{r}
map1 <- filter(map, Study == "Expt1")
otu1 <- otu[,colnames(otu) %in% map1$SampleID]

dds1 <- DESeqDataSetFromMatrix(otu1,
                                colData = map1,
                                design = ~ Soil + Inoculum)


vsd1 <- varianceStabilizingTransformation(dds1, blind = FALSE) %>% assay
vsd1[vsd1 < 0.0] <- 0.0

dist1 <- vegdist(t(vsd1))
adonis(dist1 ~ History + Soil2, map1)

pcoa.axes1 <- pcoa_axes(as.matrix(dist1), map1)
pcoa.eigval1 <- pcoa_eigval(as.matrix(dist1), map1)

expt1_pcoa <- ggplot(pcoa.axes1, aes(Axis.1, Axis.2, color = Soil2, shape = paste(Study, Inoculum, sep = " / "))) +
  geom_point(size = 3, stroke = 2) +
  scale_shape_manual(values = c(1,16), name = "Study / Treatment") +
  scale_color_manual(values = soil.pal[c(1,2,3,6)], name = "Soil") +
  xlab(paste("PCo1 (", pcoa.eigval1$Eigval[1], "%)", sep = "")) +
  ylab(paste("PCo2 (", pcoa.eigval1$Eigval[2], "%)", sep = "")) +
  theme(text = element_text(size = 12),
        panel.background = element_rect(fill = "white"),
        panel.border = element_rect(fill = NA),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank())
```

```{r}
dd.row1 <- as.dendrogram(hclust(as.dist(dist1), method = "average"))
ddata_x1 <- dendro_data(dd.row1)

dendro.p1 <- ggplot(segment(ddata_x1)) +
  geom_segment(aes(x=x, y=-y, xend=xend, yend=-yend))

labs1 <- label(ddata_x1) %>% 
  inner_join(map1, by = c("label" = "SampleID"))

phy.ra1 <- otu.ra.tidy %>% 
  inner_join(tax) %>% 
  group_by(SampleID, PhyClass2) %>%
  summarise(Total = sum(Count)) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = (Total/sum(Total))) %>% 
  inner_join(map1, by = "SampleID") %>% 
  select(SampleID, PhyClass2, RelAb)

phy.ra1 <- inner_join(labs1, phy.ra1,  by = c("label" = "SampleID"))

dendro.p1 + 
  geom_bar(data = phy.ra1, 
           aes(x=x, y=RelAb, fill = PhyClass2), stat = "identity") +
  geom_point(data=labs1,
             aes(x=x, y=-.025, colour=Soil2, shape = paste(Study, Inoculum, sep = " / ")), size = 4, stroke = 1) +
  scale_color_manual(values = soil.pal, name = "Soil") +
  scale_fill_manual(values = phy.pal) +
  scale_shape_manual(values = c(1,16), name = "Study / Treatment") +
  coord_flip(expand = T) +
  theme_void() +
  theme(text = element_text(size = 15))
```

Visaulize them separately 
First experiment
```{r}
map2 <- filter(map, Study == "Expt2")
otu2 <- otu[,colnames(otu) %in% map2$SampleID]

dds2 <- DESeqDataSetFromMatrix(otu2,
                                colData = map2,
                                design = ~ Soil + Inoculum)


vsd2 <- varianceStabilizingTransformation(dds2, blind = FALSE) %>% assay
vsd2[vsd2 < 0.0] <- 0.0



dist2 <- vegdist(t(vsd2))
adonis(dist2 ~ History + Soil2, map2)

pcoa.axes2 <- pcoa_axes(as.matrix(dist2), map2)
pcoa.eigval2 <- pcoa_eigval(as.matrix(dist2), map2)

expt2_pcoa <- ggplot(pcoa.axes2, aes(Axis.1, Axis.2, color = Soil2, shape = paste(Study, Inoculum, sep = " / "))) +
  geom_point(size = 3, stroke = 2) +
  scale_shape_manual(values = c(1,16), name = "Study / Treatment") +
  scale_color_manual(values = soil.pal, name = "Soil") +
  xlab(paste("PCo1 (", pcoa.eigval2$Eigval[1], "%)", sep = "")) +
  ylab(paste("PCo2 (", pcoa.eigval2$Eigval[2], "%)", sep = "")) +
  theme(text = element_text(size = 12),
        panel.background = element_rect(fill = "white"),
        panel.border = element_rect(fill = NA),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank())
```

```{r}
dd.row2 <- as.dendrogram(hclust(as.dist(dist2), method = "average"))
ddata_x2 <- dendro_data(dd.row2)

dendro.p2 <- ggplot(segment(ddata_x2)) +
  geom_segment(aes(x=x, y=-y, xend=xend, yend=-yend))

labs2 <- label(ddata_x2) %>% 
  inner_join(map2, by = c("label" = "SampleID"))

phy.ra2 <- otu.ra.tidy %>% 
  inner_join(tax) %>% 
  group_by(SampleID, PhyClass2) %>%
  summarise(Total = sum(Count)) %>% 
  group_by(SampleID) %>% 
  mutate(RelAb = (Total/sum(Total))) %>% 
  inner_join(map2, by = "SampleID") %>% 
  select(SampleID, PhyClass2, RelAb)

phy.ra2 <- inner_join(labs2, phy.ra2,  by = c("label" = "SampleID"))

dendro.p2 + 
  geom_bar(data = phy.ra2, 
           aes(x=x, y=RelAb, fill = PhyClass2), stat = "identity") +
  geom_point(data=labs2,
             aes(x=x, y=-.025, colour=Soil2, shape = paste(Study, Inoculum, sep = " / ")), size = 3, stroke = 1) +
  scale_color_manual(values = c("#08519C", "#54278F", "#9ECAE1", "#9E9AC8", "gold", "#DADAEB"), name = "Soil") +
  scale_fill_manual(values = c("black", RColorBrewer::brewer.pal(11, "Spectral")[c(9,2,11)], RColorBrewer::brewer.pal(6, "Greys")[-1])) +
  scale_shape_manual(values = c(1,16), name = "Study / Treatment") +
  coord_flip(expand = T) +
  theme_void() +
  theme(text = element_text(size = 15))
```

Only live inoculated seedlings
```{r}
map1_live <- filter(map, Study == "Expt1" & Inoculum == "SoilMB")
otu1_live <- otu[,colnames(otu) %in% map1_live$SampleID]

dds1_live <- DESeqDataSetFromMatrix(otu1_live,
                                colData = map1_live,
                                design = ~ Soil)


vsd1_live <- varianceStabilizingTransformation(dds1_live, blind = FALSE) %>% assay
vsd1_live[vsd1_live < 0.0] <- 0.0



dist1_live <- vegdist(t(vsd1_live))
tidy(adonis(dist1_live ~ History + Soil, map1_live)$aov.tab)

pcoa.axes1_live <- pcoa_axes(as.matrix(dist1_live), map1_live)
pcoa.eigval1_live <- pcoa_eigval(as.matrix(dist1_live), map1_live)

exp1_pcoa_live <- ggplot(pcoa.axes1_live, aes(Axis.1, Axis.2, color = Soil2, shape = paste(Study, Inoculum, sep = " / "))) +
  geom_point(size = 3, stroke = 2) +
  scale_shape_manual(values = c(16), name = "Study / Treatment") +
  scale_color_manual(values = soil.pal, name = "Soil") +
  xlab(paste("PCo1 (", pcoa.eigval1_live$Eigval[1], "%)", sep = "")) +
  ylab(paste("PCo2 (", pcoa.eigval1_live$Eigval[2], "%)", sep = "")) +
  theme(text = element_text(size = 12),
        panel.background = element_rect(fill = "white"),
        panel.border = element_rect(fill = NA),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank())
```

```{r}
map2_live <- filter(map, Study == "Expt2" & Inoculum == "SoilMB")
otu2_live <- otu[,colnames(otu) %in% map2_live$SampleID]

dds2_live <- DESeqDataSetFromMatrix(otu2_live,
                                colData = map2_live,
                                design = ~ Soil)


vsd2_live <- varianceStabilizingTransformation(dds2_live, blind = FALSE) %>% assay
vsd2_live[vsd2_live < 0.0] <- 0.0



dist2_live <- vegdist(t(vsd2_live))
tidy(adonis(dist2_live ~ History + Soil, map2_live)$aov.tab)

pcoa.axes2_live <- pcoa_axes(as.matrix(dist2_live), map2_live)
pcoa.eigval2_live <- pcoa_eigval(as.matrix(dist2_live), map2_live)

exp2_pcoa_live <- ggplot(pcoa.axes2_live, aes(Axis.1, Axis.2, color = Soil2, shape = paste(Study, Inoculum, sep = " / "))) +
  geom_point(size = 3, stroke = 2) +
  scale_shape_manual(values = c(16), name = "Study / Treatment") +
  scale_color_manual(values = soil.pal, name = "Soil") +
  xlab(paste("PCo1 (", pcoa.eigval2_live$Eigval[1], "%)", sep = "")) +
  ylab(paste("PCo2 (", pcoa.eigval2_live$Eigval[2], "%)", sep = "")) +
  theme(text = element_text(size = 12),
        panel.background = element_rect(fill = "white"),
        panel.border = element_rect(fill = NA),
        axis.ticks.x = element_blank(),
        axis.ticks.y = element_blank())
```



